---
title: "PA1_template"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load data

```{r cache=TRUE}
data<-read.csv("C:/Users/Zeeshan/Desktop/coursera/data1/activity.csv")
head(data)
library(ggplot2)
library(mice)
```

##What is mean total number of steps taken per day?

removing NA from data frame

```{r cache=TRUE}
data.wo.na <- data[complete.cases(data),]
head(data.wo.na)
```

taking sum of no. of steps taken per day

```{r cache=TRUE}
date.step<-aggregate(data.wo.na[, 1], list(data.wo.na$date), sum)
colnames(date.step)<-c("date","steps")
```

1. plotting histogram of date and no of steps

```{r cache=TRUE}
 ggplot(date.step,aes(x=date, y=steps)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

2. Calculate and report the mean and median total number of steps taken per day

```{r cache=TRUE}
Mean <- mean(date.step$steps)
Mean
Median <- median(date.step$steps)
Median
```

## What is the average daily activity pattern?

finding average no of steps taken per interval across days

```{r cache=TRUE}
averageStepsPerTimeBlock<-aggregate(data$steps, list(data$interval), mean, na.rm=T)
colnames(averageStepsPerTimeBlock)<-c("interval","meansteps")
head(averageStepsPerTimeBlock)
```

1. Make a time series plot

```{r cache=TRUE}
ggplot(data=averageStepsPerTimeBlock, aes(x=interval, y=meansteps)) +geom_line() + xlab("5-minute interval") +     ylab("average number of steps taken")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r cache=TRUE}
mostSteps <- averageStepsPerTimeBlock[which.max(averageStepsPerTimeBlock$meansteps),]
```

most steps
```{r cache=TRUE}
mostSteps
```


##Imputing missing values

1. Calculate and report the total number of missing values in the dataset

```{r cache=TRUE}
numMissingValues <- length(which(is.na(data$steps)))
```

Number of missing values
```{r cache=TRUE}
numMissingValues
```

2. Devise a strategy for filling in all of the missing values in the dataset.

* using MICE package to impute NA by predictive mean matching (pmm) 

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r cache=TRUE, results='hide'}
data1 <- data
tempdata <- mice(data1, m=5, maxit=50, meth='pmm', seed=500)
dataimputed <- complete(tempdata,1)
```

4. Make a histogram of the total number of steps taken each day

```{r cache=TRUE}
 ggplot(dataimputed,aes(x=date, y=steps)) + geom_bar(stat = "identity") + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

... and Calculate and report the mean and median total number of steps taken per day

```{r cache=TRUE}
meansteps<-mean(dataimputed$steps)
meansteps
mediansteps<-median(dataimputed$steps)
mediansteps
```


##Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r cache=TRUE}
dataimputed$datetype <-  ifelse(as.POSIXlt(dataimputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
head(dataimputed)
```


2. Make a panel plot containing a time series plot

```{r cache=TRUE}
averagedActivityDataImputed <- aggregate(steps ~ interval + datetype, data=dataimputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) +    geom_line() + facet_grid(datetype ~ .) + xlab("5-minute interval") + ylab("avarage number of steps")
```